---
title: "Reproducible Country Report"
author: "Laurie Baker"
date: "30/07/2020"
output: 
  html_document:
    theme: cosmo
    highlight: haddock
    number_sections: TRUE
params:
  country: Japan
  continent: Asia #Step 4: Adding another parameter
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, echo = FALSE, message = FALSE}

## Step 2 (changing the code chunk settings)

#install.packages("tidyverse")
library(tidyverse)

```

```{r Data Preparation, echo = FALSE, message = FALSE, warning = FALSE}

source("scripts/preprocessing.R") ## Step 1 Moving out the preprocessing to another script. 

# Country of Interest
gapminder_country <- gapminder_post %>%
  filter(country == params$country,
         year > 1980 & year <= 2010)

# Wider Continent
gapminder_continent <- gapminder_post %>%
  filter(continent == params$continent,
         year > 1980 & year <= 2010)

# Whole world
gapminder_sub <- gapminder_post %>%
  filter(year > 1980 & year <= 2010)

```


# Life expectancy in `r params$country`


```{r lifeExp-plot, echo = FALSE}

ggplot() +
  geom_line(
    data = gapminder_sub,
    aes(
      x = year,
      y = lifeExp,
      group = country,
      color = continent
    ),
    alpha = 0.2,
    lwd = 0.75
  ) +
  geom_line(
    data = gapminder_country,
    aes(x = year,
        y = lifeExp,
        color = continent
    ),
    lwd = 2
    ) +
  labs(
    x = "Year",
    y = "Life Expectancy",
    title = paste(
      "Life Expectancy over time in",
      params$country,
      "from 1980 to 2010",
      sep = " "
    ),
    color = "Continent"
  ) +
  geom_label(
    data = gapminder_country %>%
      slice(which.min(year)),
    aes(
      x = year, 
      y = lifeExp, 
      label = country
    ),
    nudge_x = 2,
    nudge_y = -2.5
  )

```


```{r average_lifeExp, echo = FALSE}

average_lifeExp <- gapminder_country %>%
  summarise(average_lifeExp = round(mean(lifeExp), 2))

```

```{r min_lifeExp, echo = FALSE}

## Step 3 Calculating another summary statistic

min_lifeExp <- gapminder_country %>%
  summarise(min_lifeExp = round(min(lifeExp), 2))

```

```{r max_lifeExp, echo = FALSE}

## Step 3 Calculating another summary statistic

max_lifeExp <- gapminder_country %>%
  summarise(max_lifeExp = round(max(lifeExp), 2))

```

The average life expectancy in `r params$country` between 1980 and 2010 was `r average_lifeExp`. The minimum life expectancy in `r params$country` between 1980 and 2010 was `r min_lifeExp`. The maximum life expectancy in `r params$country` between 1980 and 2010 was `r max_lifeExp`. 

# Life expectancy in the `r params$continent`

```{r Prepping the data for the life expectancy plot, echo = FALSE, message = FALSE, warning = FALSE}


# Summarising the mean life expectancy to plot
lifeExp_country <- gapminder_continent %>%
  group_by(country) %>%
  summarise(mean_lifeExp = mean(lifeExp))

# Get coordinates for country label

country_coords <- world_map_data %>%
  filter(country %in% params$country) %>%
  group_by(country) %>%
  summarise(mean_long = mean(long), mean_lat = mean(lat))

# Combine world geographic data with the gapminder dataset

continent_map_data <- world_map_data %>%
  filter(continent == params$continent) %>%
  left_join(lifeExp_country, by = "country")
```

```{r Creating the life expectancy plot, echo = FALSE, message = FALSE, warning = FALSE}

ggplot(data = continent_map_data) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = group,
    fill = mean_lifeExp
  )) +
  coord_quickmap() +
  geom_point(data = country_coords, 
             aes(
               x = mean_long, 
               y = mean_lat
               )
             ) +
  geom_text(
    data = country_coords,
    aes(
      x = mean_long, 
      y = mean_lat, 
      label = country
      ),
    nudge_x = 6,
    nudge_y = 3.5
  ) +
  ggtitle("Life Expectancy across the Americas 1980-2010") +
  guides(
    fill = guide_legend(
      title = "Average Life Expectancy", 
      order = 1
      )
    ) +
  theme_void()
```

<!-- Step 2 Adding a link and a reference -->

### About the data

The data comes from the [gapminder](http://www.gapminder.org/data/) dataset from the [`gapminder` package](https://github.com/jennybc/gapminder) by Jenny Bryan.